Co-moving stars in \textsl{Gaia DR1}:\\ Abundance of very wide separation co-moving pairs

### 2016-11-21 paper/ms.tex
More edits to text and figures
- change row order of the likelihood ratio figure
- change angle origin and orientation of pie plot
- consistent usage of "connected component"
- consistent use of symbols
- more emphasis on the novel results

\title{Co-moving stars in \textsl{Gaia DR1}}
\author{Semyeong Oh\altaffilmark{\pu,\lead},
        Adrian M. Price-Whelan\altaffilmark{\pu},
        David W. Hogg\altaffilmark{\ccpp,\mpia,\cca},
        Timothy D. Morton\altaffilmark{\pu},
        David N. Spergel\altaffilmark{\pu,\cca}
}

@@

\begin{abstract}
% Context
The primary sample of the \gaia\ \textsl{First Data Release} is the brand-new
\textsl{Tycho-Gaia Astrometric Solution} (\tgas). The precision and size of this sample
presents an opportunity to find new binary stars and moving groups, especially
rare binaries or sparsely populated groups.
% Aims
Here, we look for high-confidence co-moving pairs of stars using this data.
% Methods
We use a justified marginalized likelihood ratio test to separate
pairs of stars with similar three-space velocities from
those consistent with being drawn independently from the field
population.  Although we perform some visualizations using a
(bias-corrected) inverse parallax as a distance, the likelihood test
works in the observable space and uses the \gaia\ noise model
responsibly.
% Results
We find pairs of co-moving stars out to very wide
separations up to 10~pc, the limit of our search.
Some of these pairs form larger networks through mutual co-moving neighbors,
many of which correspond to the known open clusters, and stars in OB associations.
However, there still is a surprisingly large number
of very wide ($>3$~pc) separation pairs of stars that are co-moving.
We present a catalog of high-confidence co-moving pairs of stars in \tgas\
with auxiliary information,
and discuss its potential usage
for making dynamical inferences, and testing stellar models and chemical
abundance measurements.

@@

At small separations ($0.001-1$~pc), they are wide binaries (and
multiples) that are either weakly gravitationally bound or
slowly separating.
Because they have low binding energies, a sample of wide binaries is valuable
for investigating the Galactic dynamical environment.
These systems must have survived both their dynamic birth environment and 
have avoided tidal destruction along their orbit.
Thus, the statistical properties of wide binaries provide a window into both the star formation
process and the Galactic dynamics
including Galactic tides, and other massive perturbers such as molecular clouds
and MACHOs \citep{Yoo:2004aa,Allen:2014aa}.
Wide binaries are also good testbeds for stellar evolution models.
For example, main sequence-red giant wide binaries are especially valuable in
testing consistency between stellar atmosphere models as the
two stars were likely born at the same time with the same chemical compositions,
but evolved independently because of their wide separation.
Finally, calibration of stellar parameters of low-mass stars (e.g., M dwarfs),
which dominate the stellar content of the Galaxy by number, can benefit from
a larger sample of widely separated binaries containing a low mass star and a
much brighter FGK star whose stellar parameters are easier to measure
\citep[e.g.,][]{Rojas-Ayala:2012aa}.

At larger separation ($\gtrsim 1$~pc), co-moving stars are likely members of
moving groups, associations, and star clusters.
The origin of moving groups is still under active debate: are they remnants of a coeval
star formation event with similar chemical composition? Or are they formed by
dynamical effects of non-axisymmetric features of the Galaxy such as spirals
and bars? With the recent improvements in measuring chemical abundances
using high and low resolution spectroscopy, we can now start to explore these questions
in unprecedented width (statistics) and depth (detail) \todo{What do you mean by this?}.
The dynamics of cluster dissolution provides important clues to understanding
the star formation history and the dynamical evolution of the Milky Way.
In the halo, we know of more than 20 disrupting globular clusters and dwarf galaxies
(see, e.g., \citealt{Grillmair:2016} for a summary of known streams).
These tidal streams are modeled to infer the parameters of the Galactic
potential (e.g., \citealt{Kupper:2015}).
Similar processes are at work with star clusters in the disk.
However, the dynamical time is much shorter, and the dynamics will be much more
complex because of the existence of other perturbers in the disk.

@@

The primary data set used in this \documentname\ is the Tycho-Gaia Astrometric
Solution (\tgas), released as a part of Data Release 1 (\dr) of the Gaia mission
\citep{Gaia-Collaboration:2016aa,Lindegren:2016aa}.
\tgas\ contains astrometric measurements (sky position,
parallax, and proper motions) and associated covariance matrices for a large
fraction of the \project{Tycho-2} catalog \citep{2000A&A...355L..27H} with median
astrometric precision comparable to that of the \project{Hipparcos} catalog
\citep[$\approx 0.3~{\rm mas}$;][]{2007ASSL..350.....V}. In terms of parallax
signal-to-noise ($\snr = \varpi/\sigma_\varpi$), the \tgas\ catalog contains
42385 high-precision stars with $\snr > 32$.

We construct an initial sample of star pairs to search for co-moving pairs as
follows.
We first apply a global parallax signal-to-noise cut, $\snr > 8$,  to \tgas,
which leaves 619,618 stars.
For each surviving star, we then establish an initial sample of possible
co-moving partners by selecting all other stars within 10 pc with a difference
in tangential velocity below a threshold, $\absdvtan < 10$~\kms.
We ultimately build a statistical model that incorporates the covariances of the
data, but for this position cut and for visualizations we use a point-estimate
of the distance by applying a correction for the Lutz-Kelker bias
(\citealt{Lutz:1973aa}):
\begin{equation}
  \hat{d} = 1000 \, \left[\frac{\varpi}{2} \,
    \left(1 + \sqrt{1 - \frac{16}{\snr^2}} \right) \right]^{-1} \, {\rm pc}
    \label{eq:dist}
\end{equation}
where $\varpi$ is the parallax in mas.
An estimate for the difference in tangential velocity between two stars is,
then,
\begin{equation}
  \absdvtan = |\hat{d}_1 \vec\mu_1 - \hat{d}_2 \vec\mu_2|
\end{equation}
where $\vec\mu = (\mu_\alpha, \mu_\delta)$.

@@

\begin{itemize}
  \item At small separations ($<1$~pc), there is a population of pairs with
    very small tangential velocity difference ($<2$~km/s). Given that these
    pairs are very close in both position and proper motion, it is highly
    probable that they are actually co-moving wide binaries.

  \item  A sample of co-moving stars also include stars that are part of OB associations,
    moving groups, and open clusters.
    These astrophysical objects may be detected as a network of co-moving pairs.
    As the pair separation increases, the nature of co-moving pairs
    will change from binaries to those related to these larger objects,
    which generally subtend a larger angle in sky.
    Since the proper motions of two stars with the same 3D velocity
    are essentially projections of this velocity onto sphere at
    two different viewing angles,
    the larger the difference in viewing angles is, the larger the difference in tangential
    velocities will be.
    Due to this projection effect, a population of genuine co-moving pairs
    will extend to larger \absdvtan\ at larger separation.
    This indeed can be seen in Figure~\ref{fig:dv-sep} as an over-density
    in the lower right corner that gets thinner as \absdvtan\ increases.

@@

The abundance of pairs of stars with small velocity difference in
\figname~\ref{fig:dv-sep} suggests that there are a
significant number of co-moving stars in the \tgas\ data at a range
of separations.
Here we develop a method to select high-confidence co-moving
stars that properly incorporates the complex uncertainties associated with the
\gaia\ data. We make the following assumptions in order to construct a
statistical model (a likelihood function with explicit priors on our
parameters):
\begin{itemize}
  \item We assume that the uncertainties in the data---parallax, $\varpi$, and
    two proper motion components, $\propm = (\begin{array}[t]{c c} \mu_\alpha &
    \mu_\delta\end{array})^\mathsf{T}$---are Gaussian with known covariances
    $\mat{C}$. The values of covariances are provided as part of the \gaia\ data
    \citep{Lindegren:2012aa,Lindegren:2016aa}.
  \item We assume that the 3-space velocities of a given pair of stars
    $(\vec{v}_i, \vec{v}_j)$ in the \tgas\ sample (relative to the solar system
    barycenter) are either (1) co-moving with velocity drawn from a velocity
    prior $p(\vec{v})$ and  velocity difference drawn from a zero-mean Gaussian
    with velocity dispersion $s$, or (2) individually drawn from the same
    velocity prior $p(\vec{v})$.
\end{itemize}

@@

\begin{figure*}[htbp]
  \begin{center}
    \includegraphics[width=\textwidth]{figures/likelihoodratios.pdf}
  \end{center}
  \caption{%
    Justification of the likelihood ratio cut.
    On the top row, we show how the distribution of pairs on \absdvtan\ vs. separation changes
    in slices of $\ln \mathcal{L}_1 /\mathcal{L}_2$ (sequentially color-coded in blue).
    On the bottom row, we show how the distribution of $\ln \mathcal{L}_1 /\mathcal{L}_2$
    changes in slices of \absdvtan\ (sequentially color-coded in red).
    \label{fig:likelihoodratios}}
    \vspace{1em}
\end{figure*}

Figure~\ref{fig:likelihoodratios} shows the likelihood ratios for all
$\sim 271$k initial pairs examined.
As discussed in \sectionname~\ref{sec:data}, we expect a correlation between likelihood
ratios of pairs, and their distribution in \absdvtan\ vs separation plane.
Specifically, as we sweep through from small \absdvtan\ to large, we expect
the population of pairs to change from genuinely co-moving to random.
This becomes clear when we look at the distribution of likelihood ratios for
slices of \absdvtan\ centered at 0.5, 5, and 9.5~\kms.
Pairs with $|\absdvtan -0.5|<0.5$ are most likely actual co-moving pairs, and
their likelihood distribution is narrowly peaked above 6.
The distribution peaks at lower values and gets broader as \absdvtan\ increases,
and the number of random pairs increasingly dominate.
On the right panel of Figure~\ref{fig:likelihoodratios}, we show
how the distribution of pairs on \absdvtan\ vs separation plane changes
with decreasing $\ln \mathcal{L}_1 /\mathcal{L}_2$ ratios.
This is in agreement with our discussion in \sectionname~\ref{sec:data}.
Finally, as a test,
we compute the likelihood ratios for 200,000 random pairs of stars with the same
parallax signal-to-noise ratio cut as candidate pairs ($S/N > 8$).
Shown as gray shaded histogram in Figure~\ref{fig:likelihoodratios},
this distribution peaks at a much lower value ($\approx 2$),
and is clearly separated from highly probable co-moving pairs.

Based on these comparisons, we select pairs with
$\ln \mathcal{L}_1 /\mathcal{L}_2 > 6$ as co-moving pair candidates.
Out of 271,232 candidate pairs examined, 13,058 pairs (4.8\%)
satisfy this condition.

@@

\begin{figure*}[htbp]
  \begin{center}
    \includegraphics[width=\textwidth]{figures/glon_d_pie.pdf}
  \end{center}
  \caption{%
    Panoramic view of co-moving pairs of stars. This is a cylindrical
    projection of onto the Galactic plane. The angle is the Galactic
    longitude, and the radius is $d\cos b$ in pc where $d$ is the point-estimate
    of distance using Equation~\ref{eq:dist}.
    Pairs in connected components of sizes less than 5 are connected by gray lines.
    Pair in connected components of size $\geq 5$ are plotted with a unique color per network,
    and their nodes (stars) are highlighted with small black circles.
    We also show the positions of known Milky Way star clusters
    from \citet{Kharchenko:2016aa} as empty purple circles,
    and stars in OB associations from \citet{de-Zeeuw:1999aa} as
    gray circles.
    \label{fig:glon_d_pairlines}}
\end{figure*}

Once we have identified candidate co-moving star pairs from the initial pair
sample, these pairs form an undirected graph where stars are nodes, and edges
between the nodes exist for co-moving pairs of stars.
A star may have multiple co-moving neighbors, and two stars may either
be directly or indirectly connected by a set of edges.
We divide the network into connected components\footnote{A connected component of an undirected graph $G$
is a subgraph of $G$ in which any two nodes are connected to each other by a path.},
and show the distribution of their sizes in Figure~\ref{fig:hist_ccsize}.
The most common are size 2 networks, which mean mutually exclusive co-moving
pairs. However, it is clear that there are many aggregates of co-moving stars
discovered by looking for co-moving pairs.
These aggregates are likely moving groups, OB associations, or star clusters.
There are 4555 connected components among 10,606 unique stars.
The maximum size of connected components is 151.
We show this largest network along with four randomly selected
examples of networks with sizes greater than 4 in Figure~\ref{fig:graphviz_examples}.
The largest network corresponds to the Pleiades open cluster.
One particular star in this network, TYC 1799-272-1, has 94 directly connected
co-moving neighbors, and is located at the center of the network.

We show the distribution of co-moving pairs in galactic longitude and
distance in Figure~\ref{fig:glon_d_pairlines}.
The connection between known co-moving structures and the networks of co-moving
pairs found in this work becomes immediately clear when
we overplot the positions of known Milky Way star clusters \citep{Kharchenko:2016aa},
and stars in OB associations \citep{de-Zeeuw:1999aa}.
Many of the larger networks are clearly associated with known star clusters.
The most prominent is Melotte 111 at
$(l,d)\approx (220\arcdeg, 87~\textrm{pc})$ otherwise known as the Coma star cluster.
Clumps of co-moving pairs at $(l,d)\approx (300-360\arcdeg, 100-200~\textrm{pc})$
seem to strongly correlate with the locations of OB associations
Upper Scorpius, Upper Centaurus Lupus, and Lower Centaurus Crux
\citep{de-Zeeuw:1999aa}.

Ultimately, any candidate co-moving pair found in this work
needs to be verified using radial velocities.
Here, we use 210,368 cross-matches of \tgas\ with
the Radial Velocity Experiment (\rave) to
assess the false positive rates of our selection.
We have 283 pairs with both stars matched with \rave.
Figure~\ref{fig:raverv} shows the difference in radial velocities
between the two stars in a pair, $\Delta v_r$, as a function of their physical separation.
We show $\Delta v_r$ in units of $\sigma_{\Delta v_r}$ which we estimate
as the quadrature sum of $\sigma _{v_r}$ for each star.
The fraction of pairs with good agreement in radial velocity decreases with
increasing separation.
This, after all, is not surprising because we are only using
2D velocity information (proper motions) with errors.
However, the contamination becomes significant only at $>1$~pc
(which depends on the local stellar number density and velocity dispersion).
Given the excellent correspondence between co-moving pair networks of larger sizes
and known genuine co-moving structures (Figure~\ref{fig:glon_d_pairlines}),
we may expect that pairs in these larger networks which will often have
separations $>1$~pc to have less contamination.
We divide pairs into those mutually exclusively connected (i.e., in a network of
size 2), and those in a larger network.
We indeed find that many pairs in larger networks are at $>1$~pc, yet
the fraction of pairs that have identical radial velocities within $3\sigma$ is
higher than mutually exclusive pairs, and remains high ($>80\%$) to $\sim 10$~pc.

\begin{figure}[htbp]
  \begin{center}
    % \includegraphics[width=\linewidth]{figures/raverv.pdf}
    \includegraphics[width=.5\linewidth]{figures/raverv.pdf}
  \end{center}
  \caption{%
    Validation of candidate co-moving pairs using radial velocities from \rave.
    Top: Radial velocity differences, $\Delta v_r$, of 283 pairs with \rave\ measurements.
    $\Delta v_r$ is plotted in units of $\sigma_{\Delta v_r} = \sqrt{\sigma_{v_r,1}^2 + \sigma_{v_r,2}^2}$
    as a function of physical separation.
    We highlight $3\sigma$ limit with two horizontal lines.
    Bottom: Fraction of pairs with $\Delta v_r/\sigma_{\Delta v_r} <3$
    as a function of physical separation. We calculate the fraction with a running
    bin containing 31 data points at a time.
    The median, and the minimum and maximum separation of pairs in each bin are
    indicated with a marker, and its errorbars.
    We separate pairs in connected components of size 2 (i.e., mutually exclusively connected)
    from those in larger connected components.
    \label{fig:raverv}}
\end{figure}

We now examine the separation distribution of co-moving pairs in
Figure~\ref{fig:hist_separation}.
As expected, pairs in larger networks are mostly found with separations
larger than 1~pc.
Surprisingly, however, we also find a huge number of mutually exclusive pairs
at $>1$~pc as well. Even if we consider the increasing false positive rate at large
separations, the distribution is not significantly changed as the number of pairs
at $>1$~pc is in fact increasing much faster (as a power-law)
than the decrease due to the false positives (right panel of Figure~\ref{fig:raverv}).
The nature of these very large separation, mutually exclusive pairs,
which cannot be gravitationally bound to each other, needs further investigation.
Can they be remnants of escaped binaries that are drifting apart?
In a study of the evolution of wide binaries including the Galactic tidal field
as well as passing field stars, \citet{Jiang:2010aa} found that we expect
to find a peak at $\sim 100-300$~pc in the projected separation due to
stars that were once in a wide binary system, but are drifting apart with small
relative velocities ($\sim 0.1$~\kms).
Our search which stops on the maximum radius of 10~pc is obviously insufficient
to clearly detect this peak, but it shows such hint, and demonstrates
that such large scale phase-space correlations may now be looked for and studied within the \gaia\ data.

>>>>

\title{Co-moving stars in \textsl{Gaia DR1}:\\
  Abundance of very wide separation co-moving pairs}
\author{Semyeong Oh\altaffilmark{\pu,\lead},
        Adrian M. Price-Whelan\altaffilmark{\pu},
        David W. Hogg\altaffilmark{\ccpp,\mpia,\cca},
        Timothy D. Morton\altaffilmark{\pu},
        David N. Spergel\altaffilmark{\pu,\cca}
}

@@

\begin{abstract}
% Context
The primary sample of the \gaia\ \textsl{First Data Release} is the brand-new
\textsl{Tycho-Gaia Astrometric Solution} (\tgas). The precision and size of this sample
presents an opportunity to find new binary stars and moving groups, especially
rare binaries or sparsely populated groups.
% Aims
Here, we look for high-confidence co-moving pairs of stars using this data.
% Methods
We use a justified marginalized likelihood ratio test to separate
pairs of stars with similar 3D velocities from
those consistent with being drawn independently from the field
population.  Although we perform some visualizations using a
(bias-corrected) inverse parallax as a distance, the likelihood test
works in the observable space and uses the \gaia\ noise model
responsibly.
% Results
We find 13,085 pairs of co-moving stars among 10,606 unique stars
out to very wide separations up to 10~pc, the limit of our search.
Some of these pairs form a larger group through mutual co-moving neighbors,
many of which correspond to the known open clusters, and stars in OB associations.
However, we also discover many co-moving groups of pairs for the first time.
More surprisingly, we find a large number
of very wide ($>3$~pc) separation co-moving pairs of stars.
The number of these very wide separation pairs only increase with increasing
separation, and cannot be merely due to false positives.
We present a catalog of high-confidence co-moving pairs of stars in the \tgas\
with auxiliary information, and discuss its potential usage
for making dynamical inferences, and testing stellar models and chemical
abundance measurements.

@@

At small separations ($0.001-1$~pc), they are wide binaries (and
multiples) that are either weakly gravitationally bound or
slowly separating.
Because they have low binding energies, a sample of wide binaries is valuable
for investigating the Galactic dynamical environment.
These systems must have survived both their dynamic birth environment and 
have avoided tidal destruction along their orbit.
Thus, the statistical properties of wide binaries provide a window into both the star formation
process and the Galactic dynamics
including Galactic tides, and other massive perturbers such as molecular clouds
and MACHOs \citep{Yoo:2004aa,Allen:2014aa}.
Wide binaries are also good testbeds for stellar evolution models.
For example, main sequence-red giant wide binaries are especially valuable in
testing consistency between stellar atmosphere models as the
two stars were likely born at the same time with the same chemical compositions,
but evolved independently because of their wide separation.
Finally, calibration of stellar parameters of low-mass stars (e.g., M dwarfs),
which dominate the stellar content of the Galaxy by number, can benefit from
a larger sample of widely separated binaries containing a low mass star and a
much brighter F/G/K star whose stellar parameters are easier to measure
\citep[e.g.,][]{Rojas-Ayala:2012aa}.

At larger separation ($\gtrsim 1$~pc), co-moving stars are likely members of
moving groups, associations, and star clusters.
The origin of moving groups is still under active debate: are they remnants of a coeval
star formation event with similar chemical composition? Or are they formed by
dynamical effects of non-axisymmetric features of the Galaxy such as spirals
and bars? With the recent improvements in measuring chemical abundances
using high and low resolution spectroscopy, we can now start to explore these questions
with unprecedented statistics and in unexplored detail.
The dynamics of cluster dissolution provides important clues to understanding
the star formation history and the dynamical evolution of the Milky Way.
In the halo, we know of more than 20 disrupting globular clusters and dwarf galaxies
(``streams''; see, e.g., \citealt{Grillmair:2016} for a summary of known streams).
These tidal streams are modeled to infer the parameters of the Galactic
potential (e.g., \citealt{Kupper:2015}).
Similar processes are at work with star clusters in the disk.
However, the dynamical time is much shorter, and the dynamics will be much more
complex because of the existence of other perturbers in the disk.

@@

The primary data set used in this \documentname\ is the Tycho-Gaia Astrometric
Solution (\tgas), released as a part of Data Release 1 (\dr) of the Gaia mission
\citep{Gaia-Collaboration:2016aa,Lindegren:2016aa}.
The \tgas\ contains astrometric measurements (sky position,
parallax, and proper motions) and associated covariance matrices for a large
fraction of the \project{Tycho-2} catalog \citep{2000A&A...355L..27H} with median
astrometric precision comparable to that of the \project{Hipparcos} catalog
\citep[$\approx 0.3~{\rm mas}$;][]{2007ASSL..350.....V}. In terms of parallax
signal-to-noise ($\snr = \varpi/\sigma_\varpi$), the \tgas\ catalog contains
42385 high-precision stars with $\snr > 32$.

We construct an initial sample of star pairs to search for co-moving pairs as
follows.
We first apply a global parallax signal-to-noise cut, $\snr > 8$,  to the \tgas,
which leaves 619,618 stars.
For each surviving star, we then establish an initial sample of possible
co-moving partners by selecting all other stars within 10 pc with a difference
in tangential velocity below a threshold, $\absdvtan < 10$~\kms.
We ultimately build a statistical model that incorporates the covariances of the
data, but for these initial cuts and for visualizations we use a point-estimate
of the distance by applying a correction for the Lutz-Kelker bias
(\citealt{Lutz:1973aa}):
\begin{equation}
  \hat{d} = 1000 \, \left[\frac{\varpi}{2} \,
    \left(1 + \sqrt{1 - \frac{16}{\snr^2}} \right) \right]^{-1} \, {\rm pc}
    \label{eq:dist}
\end{equation}
where $\varpi$ is the parallax in mas.
An estimate for the difference in tangential velocity between two stars is,
then,
\begin{equation}
  \absdvtan = |\hat{d}_1 \vec\mu_1 - \hat{d}_2 \vec\mu_2|
\end{equation}
where $\vec\mu = (\mu_\alpha, \mu_\delta)$.

@@

\begin{itemize}
  \item At small separations ($<1$~pc), there is a population of pairs with
    very small tangential velocity difference ($<2$~km/s). Given that these
    pairs are very close in both 3D position and tangential velocities, it is highly
    probable that they are actually co-moving wide binaries.

  \item  A sample of co-moving stars also include stars that are part of OB associations,
    moving groups, and open clusters.
    These astrophysical objects may be detected as a network of co-moving pairs,
    sharing some mutual co-moving neighbors.
    As the pair separation increases, the nature of co-moving pairs
    will change from binaries to those related to these larger objects,
    which generally subtend a larger angle in sky.
    Since the proper motions of two stars with the same 3D velocity
    are essentially projections of this velocity onto the celestial sphere at
    two different viewing angles,
    the larger the difference in viewing angles is, the larger the difference in tangential
    velocities will be.
    Due to this projection effect, a population of genuine co-moving pairs
    will extend to larger \absdvtan\ at larger separation.
    This indeed can be seen in Figure~\ref{fig:dv-sep} as an over-density
    in the lower right corner that gets thinner as \absdvtan\ increases.

@@

The abundance of pairs of stars with small velocity difference in
\figname~\ref{fig:dv-sep} suggests that there are a
significant number of co-moving pairs in the \tgas\ data at a range
of separations.
Here we develop a method to select high-confidence co-moving
stars that properly incorporates the complex uncertainties associated with the
\gaia\ data. We make the following assumptions in order to construct a
statistical model (a likelihood function with explicit priors on our
parameters):
\begin{itemize}
  \item We assume that the uncertainties in the data---parallax, $\varpi$, and
    two proper motion components, $\propm = (\begin{array}[t]{c c} \mu_\alpha &
    \mu_\delta\end{array})^\mathsf{T}$---are Gaussian with known covariances
    $\mat{C}$. The values of covariances are provided as part of the \gaia\ data
    \citep{Lindegren:2012aa,Lindegren:2016aa}.
  \item We assume that the 3-space velocities of a given pair of stars
    $(\vec{v}_i, \vec{v}_j)$ in the \tgas\ sample (relative to the solar system
    barycenter) are either (1) co-moving with velocity drawn from a velocity
    prior $p(\vec{v})$ and  velocity difference drawn from a zero-mean Gaussian
    with velocity dispersion $s$, or (2) individually drawn from the same
    velocity prior $p(\vec{v})$.
\end{itemize}

@@

\begin{figure*}[htbp]
  \begin{center}
    \includegraphics[width=\textwidth]{figures/likelihoodratios.pdf}
  \end{center}
  \caption{%
    Justification of the likelihood ratio cut.
    On the top row, we show how the distribution of $\ln \mathcal{L}_1 /\mathcal{L}_2$
    changes (left) in slices of \absdvtan\ (sequentially color-coded in pink on the right panel).
    On the bottom row, we show how the distribution of pairs on \absdvtan\ vs. separation changes (right)
    in slices of $\ln \mathcal{L}_1 /\mathcal{L}_2$ (sequentially color-coded in blue on the left).
    The black dots (upper right) and lines (lower left) correspond to the entire initial sample of pairs.
    For comparision, we also present the likelihood ratio distribution of random pairs
    of stars in the \tgas\ with parallax $S/N>8$ in gray filled histogram.
    Based on this, we choose $\ln \mathcal{L}_1 /\mathcal{L}_2 > 6$ as our high-confidence
    candidate co-moving pairs.
    \label{fig:likelihoodratios}}
    \vspace{1em}
\end{figure*}

Figure~\ref{fig:likelihoodratios} shows the likelihood ratios for all
$\sim 271$k initial pairs examined.
As discussed in \sectionname~\ref{sec:data}, we expect a correlation between likelihood
ratios of pairs, and their distribution on \absdvtan\ vs separation plane.
Specifically, as we sweep through from small \absdvtan\ to large, we expect
the population of pairs to change from genuinely co-moving to random.
This becomes clear when we look at the distribution of the likelihood ratios of pairs
in slices of \absdvtan\ (top row of Figure~\ref{fig:likelihoodratios}).
Pairs with $\absdvtan \in (0,2.5)~\kms$ are most likely actual co-moving pairs, and
their likelihood ratio distribution is narrowly peaked at $\gtrsim 5$
(darkest pink histogram in the upper left panel of Figure~\ref{fig:likelihoodratios}).
The distribution peaks at lower values and gets broader as \absdvtan\ increases,
and the number of random pairs increasingly dominate.
On the bottom row of Figure~\ref{fig:likelihoodratios}, we show
how the distribution of pairs on \absdvtan\ vs separation plane changes
with decreasing $\ln \mathcal{L}_1 /\mathcal{L}_2$ ratios.
This is in agreement with our discussion in \sectionname~\ref{sec:data}.
Finally, as a test,
we compute the likelihood ratios for 200,000 random pairs of stars with the same
parallax signal-to-noise ratio cut as the initial sample ($S/N > 8$).
Shown as the gray filled histogram in Figure~\ref{fig:likelihoodratios},
this distribution peaks at a much lower value ($\approx 2$),
and is clearly separated from highly probable co-moving pairs.

Based on these comparisons, we select pairs with
$\ln \mathcal{L}_1 /\mathcal{L}_2 > 6$ as candidate co-moving pair.
Out of 271,232 pairs in the initial sample, 13,058 pairs (4.8\%)
satisfy this condition.

@@

\begin{figure*}[htbp]
  \begin{center}
    \includegraphics[width=\textwidth]{figures/glon_d_pie.pdf}
  \end{center}
  \caption{%
    Panoramic view of co-moving pairs of stars. This is a cylindrical
    projection of onto the Galactic plane. The angle is the Galactic
    longitude, and the radius is $\hat d\cos b$ in pc where $\hat d$ is the point-estimate
    of distance using Equation~\ref{eq:dist}.
    Pairs in connected components of sizes less than 5 are connected by gray lines.
    Pair in connected components of size $\geq 5$ are plotted with a unique random color,
    and their nodes (stars) are highlighted with small black circles.
    We also show the positions of known Milky Way star clusters
    from \citet{Kharchenko:2016aa} as purple circles,
    and stars in OB associations from \citet{de-Zeeuw:1999aa} as
    gray circles.
    Some of the larger connected components are clearly associated with known
    clusters, but we also discover quite a few for the first time.
    \label{fig:glon_d_pairlines}}
\end{figure*}

Once we have identified candidate co-moving pairs from the initial
sample, these pairs form an undirected graph where stars are nodes, and edges
between the nodes exist for co-moving pairs of stars.
A star may have multiple co-moving neighbors, and two stars may either
be directly or indirectly connected by a sequence of edges (``path'').
We divide the graph into connected components\footnote{A connected component of an undirected graph $G$
is a subgraph of $G$ in which any two nodes are connected to each other by a path.},
and show the distribution of their sizes in Figure~\ref{fig:hist_ccsize}.
The most common are connected components of size 2, which mean mutually exclusive co-moving
pairs. However, it is clear that there are many aggregates of co-moving stars
discovered by looking for co-moving pairs.
These aggregates are likely moving groups, OB associations, or star clusters.
In 13,058 co-moving pairs that we identified,
there are 4555 connected components among 10,606 unique stars.
The maximum size of the connected components is 151.
We show this largest connected component along with four other
examples of varying sizes in Figure~\ref{fig:graphviz_examples}.
The largest connected component corresponds to the Pleiades open cluster (left panel of Figure~\ref{fig:graphviz_examples})
while the one on the upper left panel of the right column of Figure~\ref{fig:graphviz_examples}
is NGC~2632, another known open cluster of the Milky Way \citep{Kharchenko:2016aa}.
%One particular star in this network, TYC 1799-272-1, has 94 directly connected
%co-moving neighbors, and is located at the center of the network.

We show the distribution of co-moving pairs in galactic longitude and
distance in Figure~\ref{fig:glon_d_pairlines}.
The connection between known co-moving structures and
the larger connected components found in this work becomes immediately clear when
we overplot the positions of known Milky Way star clusters \citep{Kharchenko:2016aa},
and stars in OB associations \citep{de-Zeeuw:1999aa}.
Many of the larger connected components are
clearly associated with known star clusters:
Melotte 22 (Pleiades) at $(l,d)\approx (167\arcdeg, 130~\textrm{pc})$,
Melotte 20 at $(l,d)\approx (147\arcdeg, 175~\textrm{pc})$,
Melotte 25 (Hyades) at $(l,d)\approx (180\arcdeg, 50~\textrm{pc})$, and
NGC 2632 (Beehive) at $(l,d)\approx (206\arcdeg, 187~\textrm{pc})$
to name a few.
Clumps of co-moving pairs at $(l,d)\approx (300-360\arcdeg, 100-200~\textrm{pc})$
seem to strongly correlate with the locations of OB associations
Upper Scorpius, Upper Centaurus Lupus, and Lower Centaurus Crux
\citep{de-Zeeuw:1999aa}.

However, there are still many larger connected components that we
discover for the first time.
If we define a condition to associate a connected component to a known cluster
as having more than 3 members within 10~pc of the nominal position of
the cluster, for the 61 connected components with size larger than 5,
we find that only 10 are associated with a cluster cataloged in
\citet{Kharchenko:2016aa}.
It is also worth noting that the positions of some of the known clusters
are offset from those of the connected components associated with them,
indicating that the \tgas\ data improves the distance estimates of these clusters.
Finally, not all known star clusters are recovered in our search.
This is primarily because of the non-uniform coverage of \tgas\ data.

Ultimately, any candidate co-moving pair found in this work
needs to be verified using radial velocities.
Here, we use 210,368 cross-matches of the \tgas\ with
the Radial Velocity Experiment (\rave) to
assess the false positive rates of our selection.
We have 283 pairs with both stars matched with \rave.
Figure~\ref{fig:raverv} shows the difference in radial velocities
between the two stars in a pair, $\Delta v_r$, as a function of their physical separation.
We show $\Delta v_r$ in units of $\sigma_{\Delta v_r}$ which we estimate
as the quadrature sum of $\sigma _{v_r}$ for each star.
The fraction of pairs with good agreement in radial velocity decreases with
increasing separation.
This, after all, is not surprising because we are only using
2D velocity information (proper motions) with errors.
However, the contamination becomes significant only at $>1$~pc
(which depends on the local stellar number density and velocity dispersion).
Given the excellent correspondence between
aggregates of co-moving pairs (connected components)
and known genuine co-moving structures (Figure~\ref{fig:glon_d_pairlines}),
we may expect that pairs in these larger connected components,
which will often have separations $>1$~pc,
to have less contamination from false positives.
We divide the co-moving pairs into those
mutually exclusively connected (i.e., in a connected component of
size 2), and those in a larger group.
We indeed find that many pairs in larger groups are at $>1$~pc, yet
the fraction of pairs that have identical radial velocities within $3\sigma$ is
higher than the mutually exclusive pairs, and remains high ($>80\%$) to $\sim 10$~pc.

\begin{figure}[htbp]
  \begin{center}
    % \includegraphics[width=\linewidth]{figures/raverv.pdf}
    \includegraphics[width=.5\linewidth]{figures/raverv.pdf}
  \end{center}
  \caption{%
    Validation of candidate co-moving pairs using radial velocities from \rave.
    Top: Radial velocity differences, $\Delta v_r$, of 283 pairs with \rave\ measurements.
    $\Delta v_r$ is plotted in units of $\sigma_{\Delta v_r} = \sqrt{\sigma_{v_r,1}^2 + \sigma_{v_r,2}^2}$
    as a function of physical separation.
    We highlight $3\sigma$ limit with two horizontal lines.
    Bottom: Fraction of pairs with $|\Delta v_r/\sigma_{\Delta v_r}| <3$
    as a function of physical separation. We calculate the fraction with a running
    bin containing 31 data points at a time.
    The median, and the minimum and maximum separation of pairs in each bin are
    indicated with a marker, and its errorbars.
    We separate pairs in connected components of size 2 (i.e., mutually exclusively connected)
    from those in larger connected components.
    \label{fig:raverv}}
\end{figure}

We now examine the separation distribution of co-moving pairs in
Figure~\ref{fig:hist_separation}.
As expected, pairs in larger connected components are mostly
found with separations larger than 1~pc.
Surprisingly, however, we also find a large number of mutually exclusive co-moving
pairs
at $>1$~pc as well. Even if we consider the increasing false positive rate at large
separations, the distribution is not significantly changed as the number of pairs
at $>1$~pc is in fact increasing much faster (as a power-law)
than the decrease due to the false positives
(bottom panel of Figure~\ref{fig:raverv}).
The nature of these very large separation, mutually exclusive pairs,
which cannot be gravitationally bound to each other, needs further investigation.
Can they be remnants of escaped binaries that are drifting apart?
In a study of the evolution of wide binaries including the Galactic tidal field
as well as passing field stars, \citet{Jiang:2010aa} found that we expect
to find a peak at $\sim 100-300$~pc in the projected separation due to
stars that were once in a wide binary system, but are drifting apart with small
relative velocities ($\sim 0.1$~\kms).
Our search which stops on the maximum radius of 10~pc is obviously insufficient
to clearly detect this peak, but it shows such hint, and demonstrates
that such large scale phase-space correlations may now be looked for and studied within the \gaia\ data.
